Compositional equivalences based on Open pNets

### 2020-03-30 2016-Weak-FH-Bisim/WeakBisim.tex
corrections

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s1\_m}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\underline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

    \smallskip
    The global states here are triples, build as the product of states of the 3 pLTS (remember the Holes have no state). The assignment
    $Post$ uses the variable $\texttt{m}$ from the action of hole $\texttt{P}$ to set the
    value of the sender state variable named $\texttt{s1\_m}$.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For reasons of readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$;  $\openrule
         {\{\}, True,
			() }
         {b1 \OTWeakarrow \tau b1}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {b1 \OTWeakarrow \tau b1}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b}1 \OTWeakarrow \tau b1} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\underline{out(m,ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).
 %     $\openrule
 %        {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
%			(\texttt{ec}\gets \texttt{ec}+n) }
 %        {b1 \OTWeakarrow {\underline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

@@

    \TODO{comment...}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=11cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:ImplWOA}
\end{figure}

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. We concentrate on transitions $WT_3$ and $WT_4$. Let us denote as $post_n$ the effect (= substitution) of the strong open transitions $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

$post_3 = (\texttt{s1\_ec}\gets 0, \texttt{s1\_m}\gets \texttt{m})$

@@

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_ec}\gets \texttt{s1\_ec}+1)$.
\medskip

So if we denote ${Post_{456}}^*$ any iteration of this loop, we get ${Post_{456}}^* = (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)$ for any $n\ge 0$, and the Post of the weak OT $WI_{3}$ is $ {Post_{456}}^*\shortotimes\ post_3 = (\texttt{s1\_msg}\gets m, \texttt{s1\_ec}\gets n), \forall n\ge 0$ and $Post(WI_{3'})$ is $ post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
        $($\forall n\ge 0$)

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
        $($\forall n\ge 0$)

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $($\forall n\ge 0$)

@@

$ WI_6a(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_ec}\gets \texttt{s1\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         $($\forall n\ge 0$)

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)}
  {b0 \OTarrow {\underline{\texttt{in(m1)}}} b1}
  $

$ WI_3(0) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m2)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m2}, \texttt{s1\_ec}\gets 0)}
  {202 \OTWeakarrow {\underline{\texttt{in(m2)}}} 100}
  $

@@

\begin{itemize}
  \item Their sets of active (non-silent) holes is the same: $J' = J_x = \{P\}$.
  \item Triple ($b1,  100,  \texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$) is in $\mathcal{R}$.
  \item The verification condition \\
    $\forall fv_{OT}. \{ \Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\!
   \left[\exists fv_{OT_x}.
  \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)\right]\}}$

@@

$\forall m1. \{True \land True\hspace{3mm} \implies
\exists m2. \\
(\texttt{[p-send(m1)]} = \texttt{[p-send(m2)]}
\land True
\land \underline{\texttt{in(m1)}} = \underline{\texttt{in(m2)}}\\
\land (\texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec})
\subst{(\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)
  \uplus (\texttt{s1\_msg}\gets \texttt{m2}, \texttt{s1\_ec}\gets 0)})\}
$

@@

  $ WS_4(n) = \openrule
         {\{\}, True, (\texttt{b1\_ec}\gets \texttt{b1\_ec}+n)}
         {b1 \OTWeakarrow {\tau} b1}
         \ \forall n\ge 0
         $

@@

         $(b1_msg = s1_msg \land b1_ec = s1_ec) \land True\hspace{3mm} \implies\\
\exists n.
( True \land \tau=\tau
\land (b1_msg = m1_msg \land b1_ec = m1_ec)  \subst{\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec}}
\uplus \subst{\texttt{b1\_ec}\gets \texttt{b1\_ec}+n})
$

@@

$ WI_4a(n) = \openrule
         {\{\}, True, 
    (\texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 220}
$

@@

$ WI_5a(n) = \openrule
         {\{\}, True, (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1+n)}
         {210 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

@@

$ WI_6a(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s2\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         $

@@

>>>>

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s1\_msg}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\underline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

    \smallskip
    The global states here are triples, build as the product of states of the 3 pLTS (remember the Holes have no state). The assignment
    $Post$ uses the variable $\texttt{m}$ from the action of hole $\texttt{P}$ to set the
    value of the sender state variable named $\texttt{s1\_msg}$.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For reasons of readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1}\OTWeakarrow \tau \text{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b1} \OTWeakarrow \tau \text{b1}} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\underline{out(m,ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).
 %     $\openrule
 %        {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
%			(\texttt{ec}\gets \texttt{ec}+n) }
 %        {b1 \OTWeakarrow {\underline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

@@

    \TODO{A lire...}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=11cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:ImplWOA}
\end{figure}
\RAB{In the above example, notations should be corrected once  figure \ref{SimpleProtCounter:ImplWOA} selected}

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. We concentrate on weak open transitions $WI_3$ and $WI_4$. Let us denote as $post_n$ the effect (= substitution) of the strong open transitions $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

$post_3 = (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets 0)$

@@

$post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_ec}\gets \texttt{s1\_ec}+1)$.
\medskip

So if we denote ${post_{456}}^*$ any iteration of this loop, we get ${Post_{456}}^* = (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)$ for any $n\ge 0$, and the Post of the weak OT $WI_{3}$ is $ {Post_{456}}^*\shortotimes\ post_3 = (\texttt{s1\_msg}\gets m, \texttt{s1\_ec}\gets n), \forall n\ge 0$ and $Post(WI_{3'})$ is $ post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n), \forall n\ge 0$.
\medskip

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
        $$\forall n\ge 0$

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
        $$\forall n\ge 0$

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $$\forall n\ge 0$

@@

$ WI_{6a}(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_ec}\gets \texttt{s1\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         $($\forall n\ge 0$)

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)}
  {\text{b0} \OTarrow {\underline{\texttt{in(m1)}}} \text{b1}}
  $

$ WI_3(0) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m2)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m2}, \texttt{s1\_ec}\gets 0)}
  {000 \OTWeakarrow {\underline{\texttt{in(m2)}}} 100}
  $

@@

\begin{itemize}
  \item Their sets of active (non-silent) holes is the same: $J' = J_x = \{\texttt{P}\}$.
  \item Triple ($\text{b1},  100,  \texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$) is in $\mathcal{R}$.
  \item The verification condition \\
    $\forall fv_{OT}. \{ \Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\!
   \left[\exists fv_{OT_x}.
  \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)\right]\}}$

@@

$\forall \texttt{m1}. \{True \land True\hspace{3mm} \implies
\exists \texttt{m2}. \\
(\texttt{[p-send(m1)]} = \texttt{[p-send(m2)]}
\land True
\land \underline{\texttt{in(m1)}} = \underline{\texttt{in(m2)}} \land 
\\
(\texttt{b1\_msg\!\! = \!\!s1\_msg} \land \texttt{b1\_ec\!\! =\!\! s1\_ec}) \subst{(\texttt{b1\_msg}\!\gets\! \texttt{m1}, \texttt{b1\_ec}\!\gets\! 0)\uplus (\texttt{s1\_msg}\!\gets\! \texttt{m2},\texttt{s1\_ec}\!\gets\! 0)})\}$

@@

  $ WS_4(n) = \openrule
         {\{\}, True, (\texttt{b1\_ec}\gets \texttt{b1\_ec}+n)}
         {\text{b1} \OTWeakarrow {\tau} \text{b1}}
         \ \forall n\ge 0
         $

@@

         $(\texttt{b1\_msg} = \texttt{s1\_msg} \land \texttt{b1\_ec} = \texttt{s1\_ec}) \land True\hspace{3mm} \implies
\exists n.(True \land \tau=\tau \land \\ (\texttt{b1\_msg} = \texttt{m1\_msg} \land \texttt{b1\_ec} = \texttt{m1\_ec})  \subst{\texttt{m1\_msg}\!\gets\! \texttt{s1\_msg}, \texttt{m1\_ec}\!\gets\! \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec}}
\uplus \subst{\texttt{b1\_ec}\gets \texttt{b1\_ec}+n})
$

@@

$ WI_{4a}(n) = \openrule
         {\{\}, True, 
    (\texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 220}
$

@@

$ WI_{5a}(n) = \openrule
         {\{\}, True, (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1+n)}
         {210 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

@@

$ WI_{6a}(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s2\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         $

@@

\RAB{Erreur sur les deux transition $WI_{7}$, $WI_{7a}$ et $WI_{7b}$}
